---
title: "Arabidopsis_220413"
author: "Sami Saarenpää"
date: '2022-04-13'
output: html_document
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.path='figs/', fig.width=14, fig.height=10, warning=FALSE, message=FALSE)
```

# Load the packages
```{r libraries-load, echo=FALSE}
suppressPackageStartupMessages({
  library(STutility)
  library(cowplot)
  library(harmony)  
  library(stringr)
  library(pheatmap)
})
```

#Metadata and infotable
```{r}
#raw_data_ST <- data.frame()

samples <- list.files(path = "Arabidopsis/10_45_45_all_samples/data/stdata", pattern = ".tsv", recursive = FALSE, full.names = T)

imgs <- list.files(path = "Arabidopsis/10_45_45_all_samples_R/imgs/", pattern = ".jpg", recursive = FALSE, full.names = T)

spotfiles <- list.files(path = "Arabidopsis/10_45_45_all_samples_R/spotfiles/", pattern = ".tsv", recursive = FALSE, full.names = T)

slide_id <- c(rep("128",6), rep("121", 7))
leaf <- c("1","1","2","1","2","2","3","3","3", "4", "4", "4", "4")
plant <- c("p1", "p1", "p1", "p1", "p1", "p1", "p2", "p2", "p2", "p2", "p2", "p2", "p2")
section_id <- c("128_A1", "128_A2", "128_B1", "128_B2", "128_C1", "128_C2", "121_A1", "121_A2", "121_B1", "121_B2", "121_C1", "121_C2", "121_D1")


infoTable <- data.frame(samples, 
                        spotfiles, 
                        imgs,
                        slide_id,
                        leaf,
                        plant,
                        section_id,
                        stringsAsFactors = F)

```

#Reading the data with STUtility and filtering
```{r, fig.width = 22, fig.height=18}

moderate_filter <- InputFromTable(infotable = infoTable, 
                                 transpose = F, 
                                 platform = "Visium",
                                 minUMICountsPerGene = 30,
                                 minUMICountsPerSpot = 10,
                                 minSpotsPerGene = 20,
                                 minGenesPerSpot = 10,
                                 disable.subset = F)

```

#Filtering of mitochondrial, chloroplast, ribosomal and non-coding genes
```{r, fig.width = 22, fig.height=18}
  mt.genes<- grep(pattern = "^ATM", x = rownames(moderate_filter), value = TRUE)
  chloro.genes <- grep(pattern = "^ATC", x = rownames(moderate_filter), value = TRUE)
  rb.Genes <- as.list(c("AT1G72370", "AT3G04770", "AT1G58380", "AT1G59359", "AT2G41840", "AT3G57490", "AT1G58684", "AT1G58983", "AT2G31610", "AT3G53870", "AT5G35530", "AT3G04840", "AT4G34670", "AT2G17360", "AT5G07090", "AT5G58420", "AT2G37270", "AT3G11940", "AT4G31700", "AT5G10360", "AT1G48830", "AT3G02560", "AT5G16130", "AT5G20290", "AT5G59240", "AT4G12160", "AT5G15200", "AT5G39850", "AT4G25740", "AT5G41520", "AT5G52650", "AT3G48930", "AT4G30800", "AT5G23740", "AT1G15930", "AT1G80800", "AT2G32060", "AT3G60770", "AT4G00100", "AT2G36160", "AT3G11510", "AT3G52580", "AT1G04270", "AT5G09490", "AT5G09500", "AT5G09510", "AT5G43640", "AT5G63070", "AT1G07770", "AT2G19720", "AT2G39590", "AT3G46040", "AT4G29430", "AT5G59850", "AT2G09990", "AT3G04230", "AT5G18380", "AT2G04390", "AT2G05220", "AT3G10610", "AT5G04800", "AT1G22780", "AT1G34030", "AT4G09800", "AT3G02080", "AT5G15520", "AT5G61170", "AT3G45030", "AT3G47370", "AT5G62300", "AT3G27450", "AT3G53890", "AT5G27700", "AT3G09680", "AT5G02960", "AT3G04920", "AT5G28060", "AT2G16360", "AT2G21580", "AT3G30740", "AT4G34555", "AT4G39200", "AT2G40510", "AT2G40590", "AT3G56340", "AT2G45710", "AT3G61110", "AT5G47930", "AT1G23410", "AT2G47100", "AT3G62250", "AT3G10090", "AT5G03850", "AT5G64140", "AT3G43980", "AT3G44010", "AT4G33865", "AT2G19750", "AT4G29390", "AT5G56670", "AT2G40010", "AT3G09200", "AT3G11250", "AT1G01100", "AT4G00810", "AT5G47700", "AT5G24510", "AT3G49460", "AT2G27720", "AT2G27710", "AT3G28500", "AT3G44590", "AT5G40040", "AT4G25890", "AT5G57290", "AT1G43170", "AT1G61580", "AT5G42445", "AT3G09630", "AT1G35200", "AT2G24730", "AT5G02870", "AT3G25520", "AT5G39740", "AT5G40130", "AT1G18540", "AT1G74060", "AT1G74050", "AT1G80750", "AT2G01250", "AT2G44120", "AT3G13580", "AT2G47610", "AT3G62870", "AT2G18020", "AT3G51190", "AT4G36130", "AT1G33120", "AT1G33140", "AT4G10450", "AT1G14320", "AT1G26910", "AT1G66580", "AT1G08360", "AT2G27530", "AT5G22440", "AT2G42740", "AT3G58700", "AT4G18730", "AT5G45775", "AT2G37190", "AT3G53430", "AT5G60670", "AT3G48130", "AT3G49010", "AT3G48960", "AT5G23900", "AT3G07110", "AT3G24830", "AT4G13170", "AT5G48760", "AT2G20450", "AT4G27090", "AT4G16720", "AT4G17390", "AT1G27400", "AT1G67430", "AT2G47570", "AT3G05590", "AT5G27850", "AT1G29970", "AT2G34480", "AT3G14600", "AT1G02780", "AT3G16780", "AT4G02230", "AT1G09590", "AT1G09486", "AT1G09690", "AT1G31355", "AT1G57660", "AT3G57820", "AT1G57860", "AT1G02830", "AT3G05560", "AT5G27770", "AT1G04480", "AT2G33370", "AT3G04400", "AT2G39460", "AT3G55280", "AT2G36620", "AT3G53020", "AT2G44860", "AT3G49910", "AT5G67510", "AT2G32220", "AT3G22230", "AT4G15000", "AT1G12960", "AT1G23290", "AT1G70600", "AT2G19730", "AT3G06700", "AT3G06680", "AT1G36240", "AT1G77940", "AT3G18740", "AT2G19740", "AT4G26230", "AT5G56710", "AT4G18100", "AT5G46430", "AT1G26880", "AT1G69620", "AT3G28900", "AT3G09500", "AT2G39390", "AT3G55170", "AT5G02610", "AT1G07070", "AT1G41880", "AT1G74270", "AT3G55750", "AT2G37600", "AT3G53740", "AT5G02450", "AT3G23390", "AT4G14320", "AT1G15250", "AT1G52300", "AT3G16080", "AT2G35180", "AT3G10950", "AT3G60245", "AT2G43460", "AT3G59540", "AT2G25210", "AT3G02190", "AT4G31985", "AT2G36170", "AT3G52590", "AT1G56045", "AT3G08520", "AT3G11120", "AT3G56020"))
  Araport11_rRNA.201606 <- read.table("Arabidopsis/Araport11_rRNA.201606.ID", quote="\"", comment.char="")
  Araport11_rRNA.201606 <- as.list(Araport11_rRNA.201606$V1)
  rb.Genes <- as.character(append(rb.Genes,Araport11_rRNA.201606))
   
  #Noncoding genes to remove
  noncoding.genes <-  read.table("Arabidopsis/Araport11_non_coding.201606.ID", quote="\"", comment.char="")
  noncoding.genes <- as.character(noncoding.genes$V1)
```

##Mt filter
```{r, fig.width= 8, fig.height=20}

mt.moderate_filter <- moderate_filter[setdiff(rownames(moderate_filter),mt.genes)]

```
##MtRb filter
```{r, fig.width= 8, fig.height=20}

mtrb.moderate_filter <- moderate_filter[setdiff(rownames(mt.moderate_filter),intersect(rownames(mt.moderate_filter), rb.Genes))]

```
##MtRbCp filter
```{r, fig.width= 8, fig.height=20}

mtrbcp.moderate_filter <- mtrb.moderate_filter[setdiff(rownames(mtrb.moderate_filter),chloro.genes)]

```
##Extra genes removed
```{r, fig.width= 8, fig.height=20}

mtrbcpnoncod.moderate_filter <- mtrbcp.moderate_filter[setdiff(rownames(mtrbcp.moderate_filter),intersect(rownames(mtrbcp.moderate_filter), noncoding.genes))]

```

##Filter low number spots and subsetting the object based one the section
```{r, fig.width= 8, fig.height=20}

mtrbcpnoncod.moderate_filter.subset <- SubsetSTData(mtrbcpnoncod.moderate_filter, expression = nFeature_RNA >= 10)

subset.list.mtrbcpnoncod.moderate_filter <- c()
for (i in 1:length(infoTable$section_id)){
subset.list.mtrbcpnoncod.moderate_filter[[i]] <- SubsetSTData(mtrbcpnoncod.moderate_filter.subset, spots = colnames(mtrbcpnoncod.moderate_filter.subset)[mtrbcpnoncod.moderate_filter.subset$section_id == infoTable$section_id[i]])
}
```

##Normalisation of each of the section separately to remove intra-section batch effects
```{r, fig.width= 8, fig.height=20}
ST_list.mtrbcpnoncod.moderate_filter <- lapply(subset.list.mtrbcpnoncod.moderate_filter, function(x){
SCTransform(x, verbose =T, return.only.var.genes = F)})
```

#Colors
```{r echo=FALSE, fig.width=10, fig.height=10}

col_vector1 <- c( "#697A21", "#E49AB0",  "#3E363F", "#B8B42D","#42CAFD", "#504850", "#605960", "#B8B42D", "#697A21", "#E49AB0", "#725752", "#05668D", "#00FDDC", "#42CAFD","#DA7422", "#504850", "#605960" ,"#00FDDC")
arabidopsis.fill <- data.frame(1:18, col_vector1)
```


##Integration of the section, PCA
```{r, fig.width= 15, fig.height=15}
#Intergation features for PrepSCTIntegration
arabidopsis.features.mtrbcpnoncod.moderate_filter <- SelectIntegrationFeatures(object.list = ST_list.mtrbcpnoncod.moderate_filter)
options(future.globals.maxSize = 8000 * 1024^2)

#PreIntegration step
ST_list.mtrbcpnoncod.moderate_filter <- PrepSCTIntegration(object.list = ST_list.mtrbcpnoncod.moderate_filter, anchor.features = arabidopsis.features.mtrbcpnoncod.moderate_filter)

#Merging the objects
mtrbcpnoncod.moderate_filter.merged <- MergeSTData(ST_list.mtrbcpnoncod.moderate_filter[[1]], ST_list.mtrbcpnoncod.moderate_filter[2:length(ST_list.mtrbcpnoncod.moderate_filter)], merge.data = TRUE)
VariableFeatures(object = mtrbcpnoncod.moderate_filter.merged) <- arabidopsis.features.mtrbcpnoncod.moderate_filter

#PCA
mtrbcpnoncod.moderate_filter.merged.PCA <- RunPCA(object = mtrbcpnoncod.moderate_filter.merged, assay = "SCT", features = arabidopsis.features.mtrbcpnoncod.moderate_filter)

```

##Normalisation with Harmony to remove batch effects between samples and UMAP visualisation
```{r, fig.width= 10, fig.height=5}
#Harmony
mtrbcpnoncod.moderate_filter.merged.PCA.harmony <- RunHarmony(mtrbcpnoncod.moderate_filter.merged.PCA, group.by.vars =c("slide_id", "plant", "leaf"), plot_convergence = TRUE, theta = c(4,4,4), assay.use = "SCT", reduction = "pca", dims.use = 1:8)

#UMAP
mtrbcpnoncod.moderate_filter.merged.PCA.harmony <- RunUMAP( object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony, dims = 1:8, assay = "SCT", reduction = "harmony")
mtrbcpnoncod.moderate_filter.merged.PCA.harmony <- FindNeighbors(object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony, assay = "SCT" ,reduction = "harmony", k.param = 15, dims = 1:8)
mtrbcpnoncod.moderate_filter.merged.PCA.harmony <- FindClusters(object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony, resolution = 0.22, random.seed = 5)

#UMAP visualisation
DimPlot(object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony, reduction = "umap", cols = col_vector1)

```

```{r, fig.width= 10, fig.height=5}

VlnPlot(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, features = c("nFeature_SCT", "nCount_SCT"), ncol = 2, pt.size = F, group.by = "leaf", log = T, same.y.lims = F, split.plot = T, split.by = "seurat_clusters") 

```


#Opacity change, masking the image
```{r, fig.width= 16, fig.height=13}

mtrbcpnoncod.moderate_filter.merged.PCA.harmony <- LoadImages(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, time.resolve = F, xdim = 900)
mtrbcpnoncod.moderate_filter.merged.PCA.harmony <- MaskImages(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, channels.use = 1, iso.blur = 3)

st.object <- GetStaffli(mtrbcpnoncod.moderate_filter.merged.PCA.harmony) # Collect Staffli object from Seurat object

# following part will modify the "raw" images
# If you want to change the opacity for "masked" or "processed" images you will need to specify this
st.object@rasterlists$masked <- lapply(st.object@rasterlists$masked, function(im) {
  im <- apply(im, 2, function(x) {
    scales::alpha(x, alpha = 0.6)
  })
  return(im)
})
mtrbcpnoncod.moderate_filter.merged.PCA.harmony@tools$Staffli <- st.object

ImagePlot(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, type = "masked")
```

##UMAP clusters on tissue
```{r, fig.width= 6, fig.height=4.5}

se <- mtrbcpnoncod.moderate_filter.merged.PCA.harmony
col_vector1 <- c("#B8B42D", "#697A21", "#DC0073", "#DA7422", "#510D0A", "#3E363F", "#E49AB0")
plots1 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots1[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[1]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots2 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots2[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[2]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots3 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots3[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[3]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots4 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots4[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[4]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots5 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots5[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[5]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots6 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots6[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[6]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots7 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots7[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[7]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots8 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots8[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[8]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots9 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots9[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[9]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots10 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots10[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[10]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots11 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots11[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[11]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots12 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots12[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[12]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots13 <- c()
for (i in 1:length(unique(se$seurat_clusters))){
  plots13[[i]] =FeatureOverlay(object = SubsetSTData(se, 
                                                   spots = colnames(se)[se$section_id == infoTable$section_id[[13]] & se$seurat_clusters == levels(se$seurat_clusters)[[i]]]),
                             features = "seurat_clusters",
                             pt.size = 1.1,
                             show.sb = F,
                             sample.label = F,
                             cols = col_vector1[[i]],
                             dark.theme = F,
                             pt.alpha = 0.8)
}

```


#Marker detection
```{r}
DefaultAssay(se) <- "RNA"
NormalizeData(se, assay = "RNA", normalization.method = "LogNormalize") 
Markers.mtrbcpnoncod.5cl.RNA.poisson <- FindAllMarkers(se, test.use = "poisson", logfc.threshold = 0.15, min.pct = 0.02, only.pos = F, random.seed = 5678)
markers <- Markers.mtrbcpnoncod.5cl.RNA.poisson$gene

```

##Marker visualisation
```{r, fig.width= 15, fig.height=14}

FeatureOverlay(object = se,
                        features = "AT1G29910",
                        spots = rownames(subset(se[[]],GetAssayData(object=se,assay="RNA",slot="counts")["AT1G29910",] >= 1)),
                        pt.size = 1.1,
                        sampleids = 1:13,
                        type = "masked",
                        cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                        pt.alpha = 1,
                        ncols = 4)

FeatureOverlay(object = se,
                        features = "AT2G30860",
                        spots = rownames(subset(se[[]],GetAssayData(object=se,assay="RNA",slot="counts")["AT2G30860",] >= 1)),
                        pt.size = 1.1,
                        sampleids = 1:13,
                        type = "masked",
                        cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                        pt.alpha = 1,
                        ncols = 4)

FeatureOverlay(object = se,
                        features = "AT2G35750",
                        spots = rownames(subset(se[[]],GetAssayData(object=se,assay="RNA",slot="counts")["AT2G35750",] >= 1)),
                        pt.size = 1.1,
                        sampleids = 1:13,
                        type = "masked",
                        cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                        pt.alpha = 1,
                        ncols = 4)

FeatureOverlay(object = se,
                        features = "AT2G31141",
                        spots = rownames(subset(se[[]],GetAssayData(object=se,assay="RNA",slot="counts")["AT2G31141",] >= 1)),
                        pt.size = 1.1,
                        sampleids = 1:13,
                        type = "masked",
                        cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                        pt.alpha = 1,
                        ncols = 4)

```

##Subset object based on the clusters
```{r, fig.width= 12, fig.height=20}
se.cl1 <- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony,spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters == levels(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters)[[1]]])

se.cl2 <- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony,spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters == levels(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters)[[2]]])

se.cl3 <- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony,spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters == levels(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters)[[3]]])

se.cl4 <- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony,spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters == levels(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters)[[4]]])

se.cl5 <- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony,spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters == levels(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$seurat_clusters)[[5]]])
```

##Number of genes and unique molecules -- Arabidopsis
```{r, fig.width= 12, fig.height=20}
DefaultAssay(mtrbcpnoncod.moderate_filter.merged.PCA.harmony) <- "RNA"
ugenes <- nrow(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)

umolecules <- sum(rowSums(mtrbcpnoncod.moderate_filter.merged.PCA.harmony))

unique_genes <- c()
unique_molecules <- c()
temp_sample <- c()
keep.genes <- c()
spots_sample <- c()
for (i in 1:length(infoTable$section_id)){
  temp_sample[[i]] <- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[i]])
  keep.genes[[i]] <- rownames(temp_sample[[i]][rowSums(temp_sample[[i]]@assays$RNA@counts) >0])
  unique_genes[[i]] <- length(keep.genes[[i]])
  temp_sample[[i]] <- temp_sample[[i]][intersect(rownames(temp_sample[[i]]), keep.genes[[i]])]
  unique_molecules[[i]] <- sum(rowSums(temp_sample[[i]]))
  spots_sample[[i]] <- ncol(temp_sample[[i]])
}

remove(temp_sample)
#Mean unique genes per sample
mean(as.numeric(unlist(unique_genes)))
#Mean unique molecules per sample
mean(as.numeric(unlist(unique_molecules)))
```

#Arabidopsis unique molecules and genes
```{r, fig.width= 12, fig.height=7}
mtrbcpnoncod.moderate_filter.merged.PCA.harmony$log.nFeature_RNA <- log10(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$nFeature_RNA + 1)
mtrbcpnoncod.moderate_filter.merged.PCA.harmony$log.nCount_RNA <- log10(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$nCount_RNA + 1)

plots <- c()
for (i in 1:13){
plots[[i]] <- FeatureOverlay(object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony,
               features = c("log.nCount_RNA", "log.nFeature_RNA"),
                                  pt.size = 1.8,
                                  sampleids = i,
                                  type = "masked",
                                  cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
               ncols = 2,
               pt.alpha = 0.5,
               pt.border = FALSE)
}
```


#Process and load the bacterial reads to the object as a new assay
```{r}
bacteria <- list.files(path = "Arabidopsis/10_45_45_all_samples/bacterial_data_210811/EXP104545/", pattern = "bacteria.csv", recursive = FALSE, full.names = T)
EXP_bacteria <- c()
for (i in 1:length(bacteria)){
  EXP_bacteria[[i]] <- read.csv(bacteria[[i]], row.names=1, sep=";", header = F)
}
colnames(EXP_bacteria[[1]]) <-lapply(lapply(EXP_bacteria[[1]][1,], function(x) paste(x,'_1')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[2]]) <-lapply(lapply(EXP_bacteria[[2]][1,], function(x) paste(x,'_2')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[3]]) <-lapply(lapply(EXP_bacteria[[3]][1,], function(x) paste(x,'_3')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[4]]) <-lapply(lapply(EXP_bacteria[[4]][1,], function(x) paste(x,'_4')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[5]]) <-lapply(lapply(EXP_bacteria[[5]][1,], function(x) paste(x,'_5')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[6]]) <-lapply(lapply(EXP_bacteria[[6]][1,], function(x) paste(x,'_6')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[7]]) <-lapply(lapply(EXP_bacteria[[7]][1,], function(x) paste(x,'_7')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[8]]) <-lapply(lapply(EXP_bacteria[[8]][1,], function(x) paste(x,'_8')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[9]]) <-lapply(lapply(EXP_bacteria[[9]][1,], function(x) paste(x,'_9')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[10]]) <-lapply(lapply(EXP_bacteria[[10]][1,], function(x) paste(x,'_10')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[11]]) <-lapply(lapply(EXP_bacteria[[11]][1,], function(x) paste(x,'_11')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[12]]) <-lapply(lapply(EXP_bacteria[[12]][1,], function(x) paste(x,'_12')), function(x) str_remove(x, " "))
colnames(EXP_bacteria[[13]]) <-lapply(lapply(EXP_bacteria[[13]][1,], function(x) paste(x,'_13')), function(x) str_remove(x, " "))

for (i in  1:length(EXP_bacteria)){
  
  EXP_bacteria[[i]] <- EXP_bacteria[[i]][-1,]
}

brns <- unique(unlist(lapply(EXP_bacteria, rownames)))
bcns <- colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)
bac_df <- data.frame(matrix(0, nrow = length(brns), ncol = length(bcns)))
rownames(bac_df) <- brns
colnames(bac_df) <- bcns
for (i in c(1:length(EXP_bacteria))){
  bac_df[rownames(EXP_bacteria[[i]]), colnames(EXP_bacteria[[i]])] = EXP_bacteria[[i]]
}
orig_s <- colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)
orig_b <- colnames(bac_df)
rmoves <- setdiff(orig_b, orig_s)
bac_df_filt <- bac_df[ , -which(names(bac_df) %in% rmoves)]
bac_df_filt[is.na(bac_df_filt)]<- 0


mtrbcpnoncod.moderate_filter.merged.PCA.harmony[['BACTERIA']] <- CreateAssayObject(counts = bac_df_filt)
DefaultAssay(mtrbcpnoncod.moderate_filter.merged.PCA.harmony) <- "BACTERIA"

```
#Process and load the fungal reads to the object as a new assay
```{r}

ITS <- list.files(path = "Arabidopsis/10_45_45_all_samples/bacterial_data_210811/EXP104545/", pattern = "ITS.csv", recursive = FALSE, full.names = T)
EXP_ITS <- c()
for (i in 1:length(ITS)){
  EXP_ITS[[i]] <- read.csv(ITS[[i]], row.names=1, sep=";", header = F)
}
colnames(EXP_ITS[[1]]) <-lapply(lapply(EXP_ITS[[1]][1,], function(x) paste(x,'_1')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[2]]) <-lapply(lapply(EXP_ITS[[2]][1,], function(x) paste(x,'_2')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[3]]) <-lapply(lapply(EXP_ITS[[3]][1,], function(x) paste(x,'_3')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[4]]) <-lapply(lapply(EXP_ITS[[4]][1,], function(x) paste(x,'_4')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[5]]) <-lapply(lapply(EXP_ITS[[5]][1,], function(x) paste(x,'_5')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[6]]) <-lapply(lapply(EXP_ITS[[6]][1,], function(x) paste(x,'_6')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[7]]) <-lapply(lapply(EXP_ITS[[7]][1,], function(x) paste(x,'_7')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[8]]) <-lapply(lapply(EXP_ITS[[8]][1,], function(x) paste(x,'_8')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[9]]) <-lapply(lapply(EXP_ITS[[9]][1,], function(x) paste(x,'_9')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[10]]) <-lapply(lapply(EXP_ITS[[10]][1,], function(x) paste(x,'_10')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[11]]) <-lapply(lapply(EXP_ITS[[11]][1,], function(x) paste(x,'_11')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[12]]) <-lapply(lapply(EXP_ITS[[12]][1,], function(x) paste(x,'_12')), function(x) str_remove(x, " "))
colnames(EXP_ITS[[13]]) <-lapply(lapply(EXP_ITS[[13]][1,], function(x) paste(x,'_13')), function(x) str_remove(x, " "))



for (i in  1:length(EXP_ITS)){
  
  EXP_ITS[[i]] <- EXP_ITS[[i]][-1,]
}

irns <- unique(unlist(lapply(EXP_ITS, rownames)))
icns <- colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)
its_df <- data.frame(matrix(0, nrow = length(irns), ncol = length(icns)))
rownames(its_df) <- irns
colnames(its_df) <- icns
for (i in c(1:length(EXP_ITS))){
  its_df[rownames(EXP_ITS[[i]]), colnames(EXP_ITS[[i]])] = EXP_ITS[[i]]
}
orig_si <- colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)
orig_bi <- colnames(its_df)
rmovesi <- setdiff(orig_bi, orig_si)
its_df_filt <- its_df[ , -which(names(its_df) %in% rmovesi)]
its_df_filt[is.na(its_df_filt)]<- 0


mtrbcpnoncod.moderate_filter.merged.PCA.harmony[['ITS']] <- CreateAssayObject(counts = its_df_filt)
DefaultAssay(mtrbcpnoncod.moderate_filter.merged.PCA.harmony) <- "ITS"

```


##Number of genes and unique molecules -- Bacteria
```{r, fig.width= 12, fig.height=20}
DefaultAssay(mtrbcpnoncod.moderate_filter.merged.PCA.harmony) <- "BACTERIA"

bacUMIs <- rowSums(GetAssayData(object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony, assay = "BACTERIA"))

bacUMIs_filt <-bacUMIs[bacUMIs>30]

ugenes_bac <- nrow(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)
umolecules_bac <- sum(rowSums(mtrbcpnoncod.moderate_filter.merged.PCA.harmony))

```


##Number of genes and unique molecules -- Fungi
```{r, fig.width= 12, fig.height=20}
DefaultAssay(mtrbcpnoncod.moderate_filter.merged.PCA.harmony) <- "ITS"

funUMIs <- rowSums(GetAssayData(object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony, assay = "ITS"))

funUMIs_filt <-funUMIs[funUMIs>30]

ugenes_its <- nrow(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)
umolecules_its <- sum(rowSums(mtrbcpnoncod.moderate_filter.merged.PCA.harmony))

```

#Microbial unique molecules and genes
```{r, fig.width= 18, fig.height=30}
mtrbcpnoncod.moderate_filter.merged.PCA.harmony$log.nFeature_BACTERIA <- log10(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$nFeature_BACTERIA + 1)
mtrbcpnoncod.moderate_filter.merged.PCA.harmony$log.nCount_BACTERIA <- log10(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$nCount_BACTERIA + 1)
mtrbcpnoncod.moderate_filter.merged.PCA.harmony$log.nFeature_ITS <- log10(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$nFeature_ITS + 1)
mtrbcpnoncod.moderate_filter.merged.PCA.harmony$log.nCount_ITS <- log10(mtrbcpnoncod.moderate_filter.merged.PCA.harmony$nCount_ITS +1 )

plots <- c()

for (i in 1:13){
plots[[i]] <- FeatureOverlay(object = mtrbcpnoncod.moderate_filter.merged.PCA.harmony,
               features = c("log.nCount_BACTERIA", "log.nFeature_BACTERIA", "log.nCount_ITS", "log.nFeature_ITS"),
                                  pt.size = 1.8,
                                  sampleids = i,
                                  type = "masked",
                                  cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
               ncols = 2, 
               pt.alpha = 0.5,
               pt.border = FALSE)
}
```

#StereoScope
```{r, fig.width= 3, fig.height=30}
#Select spots in the filtered dataset
D128A1_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[1]]]))
#Read STSC output
stsc_D128_A1<- read.table("Arabidopsis/revision/StereoScope/D128A1/W.2022-10-05161912.382856.tsv")
stsc_D128_A1_row <- rownames(stsc_D128_A1)

#Select the spots that are in the seurat object
diffRem_D128A1<- setdiff(stsc_D128_A1_row,D128A1_col)
#Select correct spots from STSC output and order them as in seurat object
stsc_D128_A1 <- stsc_D128_A1[!(row.names(stsc_D128_A1) %in% diffRem_D128A1),]
stsc_D128_A1 %>% arrange(factor(rownames(stsc_D128_A1), levels = D128A1_col))

D128A2_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[2]]]))
stsc_D128_A2<- read.table("Arabidopsis/revision/StereoScope/D128A2/W.2022-10-05161912.382856.tsv")
stsc_D128_A2_row <- rownames(stsc_D128_A2)

diffRem_D128A2<- setdiff(stsc_D128_A2_row,D128A2_col)

stsc_D128_A2 <- stsc_D128_A2[!(row.names(stsc_D128_A2) %in% diffRem_D128A2),]
stsc_D128_A2 %>% arrange(factor(rownames(stsc_D128_A2), levels = D128A2_col))

D128B1_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[3]]]))
stsc_D128_B1<- read.table("Arabidopsis/revision/StereoScope/D128B1/W.2022-10-05161912.382856.tsv")
stsc_D128_B1_row <- rownames(stsc_D128_B1)

diffRem_D128B1<- setdiff(stsc_D128_B1_row,D128B1_col)

stsc_D128_B1 <- stsc_D128_B1[!(row.names(stsc_D128_B1) %in% diffRem_D128B1),]
stsc_D128_B1 %>% arrange(factor(rownames(stsc_D128_B1), levels = D128B1_col))

D128B2_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[4]]]))
stsc_D128_B2<- read.table("Arabidopsis/revision/StereoScope/D128B2/W.2022-10-05161912.382856.tsv")
stsc_D128_B2_row <- rownames(stsc_D128_B2)

diffRem_D128B2<- setdiff(stsc_D128_B2_row,D128B2_col)

stsc_D128_B2 <- stsc_D128_B2[!(row.names(stsc_D128_B2) %in% diffRem_D128B2),]
stsc_D128_B2 %>% arrange(factor(rownames(stsc_D128_B2), levels = D128B2_col))

D128C1_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[5]]]))
stsc_D128_C1<- read.table("Arabidopsis/revision/StereoScope/D128C1/W.2022-10-05161912.382856.tsv")
stsc_D128_C1_row <- rownames(stsc_D128_C1)

diffRem_D128C1<- setdiff(stsc_D128_C1_row,D128C1_col)

stsc_D128_C1 <- stsc_D128_C1[!(row.names(stsc_D128_C1) %in% diffRem_D128C1),]
stsc_D128_C1 %>% arrange(factor(rownames(stsc_D128_C1), levels = D128C1_col))

D128C2_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[6]]]))
stsc_D128_C2<- read.table("Arabidopsis/revision/StereoScope/D128C2/W.2022-10-05161912.382856.tsv")
stsc_D128_C2_row <- rownames(stsc_D128_C2)

diffRem_D128C2<- setdiff(stsc_D128_C2_row,D128C2_col)

stsc_D128_C2 <- stsc_D128_C2[!(row.names(stsc_D128_C2) %in% diffRem_D128C2),]
stsc_D128_C2 %>% arrange(factor(rownames(stsc_D128_C2), levels = D128C2_col))

D121A1_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[7]]]))
stsc_D121_A1<- read.table("Arabidopsis/revision/StereoScope/D121A1/W.2022-10-05161912.382856.tsv")
stsc_D121_A1_row <- rownames(stsc_D121_A1)

diffRem_D121A1<- setdiff(stsc_D121_A1_row,D121A1_col)

stsc_D121_A1 <- stsc_D121_A1[!(row.names(stsc_D121_A1) %in% diffRem_D121A1),]
stsc_D121_A1 %>% arrange(factor(rownames(stsc_D121_A1), levels = D121A1_col))

D121A2_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[8]]]))
stsc_D121_A2<- read.table("Arabidopsis/revision/StereoScope/D121A2/W.2022-10-05161912.382856.tsv")
stsc_D121_A2_row <- rownames(stsc_D121_A2)

diffRem_D121A2<- setdiff(stsc_D121_A2_row,D121A2_col)

stsc_D121_A2 <- stsc_D121_A2[!(row.names(stsc_D121_A2) %in% diffRem_D121A2),]
stsc_D121_A2 %>% arrange(factor(rownames(stsc_D121_A2), levels = D121A2_col))

D121B1_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[9]]]))
stsc_D121_B1<- read.table("Arabidopsis/revision/StereoScope/D121B1/W.2022-10-05161912.382856.tsv")
stsc_D121_B1_row <- rownames(stsc_D121_B1)

diffRem_D121B1<- setdiff(stsc_D121_B1_row,D121B1_col)

stsc_D121_B1 <- stsc_D121_B1[!(row.names(stsc_D121_B1) %in% diffRem_D121B1),]
stsc_D121_B1 %>% arrange(factor(rownames(stsc_D121_B1), levels = D121B1_col))

D121B2_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[10]]]))
stsc_D121_B2<- read.table("Arabidopsis/revision/StereoScope/D121B2/W.2022-10-05161912.382856.tsv")
stsc_D121_B2_row <- rownames(stsc_D121_B2)

diffRem_D121B2<- setdiff(stsc_D121_B2_row,D121B2_col)

stsc_D121_B2 <- stsc_D121_B2[!(row.names(stsc_D121_B2) %in% diffRem_D121B2),]
stsc_D121_B2 %>% arrange(factor(rownames(stsc_D121_B2), levels = D121B2_col))


D121C1_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[11]]]))
stsc_D121_C1<- read.table("Arabidopsis/revision/StereoScope/D121C1/W.2022-10-05161912.382856.tsv")
stsc_D121_C1_row <- rownames(stsc_D121_C1)

diffRem_D121C1<- setdiff(stsc_D121_C1_row,D121C1_col)

stsc_D121_C1 <- stsc_D121_C1[!(row.names(stsc_D121_C1) %in% diffRem_D121C1),]
stsc_D121_C1 %>% arrange(factor(rownames(stsc_D121_C1), levels = D121C1_col))


D121C2_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[12]]]))
stsc_D121_C2<- read.table("Arabidopsis/revision/StereoScope/D121C2/W.2022-10-05161912.382856.tsv")
stsc_D121_C2_row <- rownames(stsc_D121_C2)

diffRem_D121C2<- setdiff(stsc_D121_C2_row,D121C2_col)

stsc_D121_C2 <- stsc_D121_C2[!(row.names(stsc_D121_C2) %in% diffRem_D121C2),]
stsc_D121_C2 %>% arrange(factor(rownames(stsc_D121_C2), levels = D121C2_col))


D121D1_col<- colnames(SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[13]]]))
stsc_D121_D1<- read.table("Arabidopsis/revision/StereoScope/D121D1/W.2022-10-05161912.382856.tsv")
stsc_D121_D1_row <- rownames(stsc_D121_D1)

diffRem_D121D1<- setdiff(stsc_D121_D1_row,D121D1_col)

stsc_D121_D1 <- stsc_D121_D1[!(row.names(stsc_D121_D1) %in% diffRem_D121D1),]
stsc_D121_D1 %>% arrange(factor(rownames(stsc_D121_D1), levels = D121D1_col))

#Merge all the dataframes
stsc <- rbind(stsc_D128_A1, stsc_D128_A2)
stsc <- rbind(stsc, stsc_D128_B1)
stsc <- rbind(stsc, stsc_D128_B2)
stsc <- rbind(stsc, stsc_D128_C1)
stsc <- rbind(stsc, stsc_D128_C2)
stsc <- rbind(stsc, stsc_D121_A1)
stsc <- rbind(stsc, stsc_D121_A2)
stsc <- rbind(stsc, stsc_D121_B1)
stsc <- rbind(stsc, stsc_D121_B2)
stsc <- rbind(stsc, stsc_D121_C1)
stsc <- rbind(stsc, stsc_D121_C2)
stsc <- rbind(stsc, stsc_D121_D1)

#Create STSC assay
mtrbcpnoncod.moderate_filter.merged.PCA.harmony[['STSC']] <- CreateAssayObject(counts = t(stsc))
DefaultAssay(mtrbcpnoncod.moderate_filter.merged.PCA.harmony) <- "STSC"
#to prevent confusion we will rename the column "nFeature_RNA" to "n_celltypes" and the column "nCount_RNA" to "sum_pp"
names(mtrbcpnoncod.moderate_filter.merged.PCA.harmony@meta.data)[18] <- "sum_pp"
names(mtrbcpnoncod.moderate_filter.merged.PCA.harmony@meta.data)[19] <- "n_celltypes"

```


For a better visual representation of the distribution of the majority of proportion values for each cell type across the tissue, the values were scaled using quantile scaling. For each section and cell type the 95th percentile of the data was considered. The values were re-scaled to the maximum value for each cell type, resulting in values between 0 and 1 within the same function. 
```{r, quantile scaling function}
scaler <- function(x, q = 0.95) {x[x > quantile(x, q)] <- max(x); scales::rescale(x)}
```

```{r, fig.height=10, fig.width=1}
#Subset per sample
D128A1<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[1]]])
D128A2<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[2]]])
D128B1<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[3]]])
D128B2<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[4]]])
D128C1<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[5]]])
D128C2<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[6]]])
D121A1<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[7]]])
D121A2<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[8]]])

D121B1<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[9]]])

D121B2<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[10]]])

D121C1<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[11]]])

D121C2<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[12]]])
D121D1<- SubsetSTData(mtrbcpnoncod.moderate_filter.merged.PCA.harmony, spots = colnames(mtrbcpnoncod.moderate_filter.merged.PCA.harmony)[mtrbcpnoncod.moderate_filter.merged.PCA.harmony$section_id == infoTable$section_id[[13]]])

#list the objects
se.l <- c(D128A1, D128A2,D128B1, D128B2, D128C1, D128C2, D121A1, D121A2, D121B1, D121B2, D121C1, D121C2, D121D1)
mat.l <- lapply(se.l, function(x){
  as.data.frame(t(as.matrix(x@assays$STSC@data)))
})
#scale the values
scale2 <- sapply(mat.l[[1]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[1]])
D128A1[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D128A1) <- "scaled2"

scale2 <- sapply(mat.l[[2]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[2]])
D128A2[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D128A2) <- "scaled2"

scale2 <- sapply(mat.l[[3]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[3]])
D128B1[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D128B1) <- "scaled2"

scale2 <- sapply(mat.l[[4]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[4]])
D128B2[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D128B2) <- "scaled2"

scale2 <- sapply(mat.l[[5]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[5]])
D128C1[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D128C1) <- "scaled2"

scale2 <- sapply(mat.l[[6]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[6]])
D128C2[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D128C2) <- "scaled2"

scale2 <- sapply(mat.l[[7]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[7]])
D121A1[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D121A1) <- "scaled2"

scale2 <- sapply(mat.l[[8]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[8]])
D121A2[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D121A2) <- "scaled2"

scale2 <- sapply(mat.l[[9]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[9]])
D121B1[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D121B1) <- "scaled2"

scale2 <- sapply(mat.l[[10]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[10]])
D121B2[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D121B2) <- "scaled2"

scale2 <- sapply(mat.l[[11]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[11]])
D121C1[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D121C1) <- "scaled2"

scale2 <- sapply(mat.l[[12]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[12]])
D121C2[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D121C2) <- "scaled2"

scale2 <- sapply(mat.l[[13]], scaler, q =0.95, USE.NAMES = TRUE)
rownames(scale2) <- rownames(mat.l[[13]])
D121D1[["scaled2"]]<-CreateAssayObject(counts = as.matrix(t(scale2)))
DefaultAssay(D121D1) <- "scaled2"

se.l <- list(D128A2,D128B1, D128B2, D128C1, D128C2, D121A1, D121A2, D121B1, D121B2, D121C1, D121C2, D121D1)
#Merge the objects back together
se.pp <- MergeSTData(D128A1, se.l)
```


#Opacity change, masking the image
```{r, fig.width= 16, fig.height=13}

se.pp <- LoadImages(se.pp, time.resolve = F, xdim = 900)
se.pp <- MaskImages(se.pp, channels.use = 1, iso.blur = 3)

st.object <- GetStaffli(se.pp) # Collect Staffli object from Seurat object

# following part will modify the "raw" images
# If you want to change the opacity for "masked" or "processed" images you will need to specify this
st.object@rasterlists$masked <- lapply(st.object@rasterlists$masked, function(im) {
  im <- apply(im, 2, function(x) {
    scales::alpha(x, alpha = 0.6)
  })
  return(im)
})
se.pp@tools$Staffli <- st.object

ImagePlot(se.pp, type = "masked")

stsc_data <- t(as.data.frame(GetAssayData(object = se.pp, assay = "STSC")))
write.csv(stsc_data, "Arabidopsis/revision/230305_semiwild_raw_stsc.csv", row.names = TRUE)

```

After each section was scaled individually the new objects with adjusted default assays for analysis can be combined again and plotted on the spot coordinates or the brightfield tissue images loaded previously if desired. 
```{r, fig.height=6, fig.width=12}

DefaultAssay(se.pp) <- "scaled2"


FeatureOverlay(object = se.pp, 
               features = c("X0","X1", "X2","X3","X4","X5","X6","X7","X8","X9","X10", "X11","X12","X13","X14","X15","X16", "X17", "X18"), 
               cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"), 
               ncols = 5,
               pt.size = 1.1,
               sampleids = 1,
               show.sb = FALSE,
               value.scale = "all")


FeatureOverlay(object = se.pp, 
               features = c("X0","X1", "X2","X3","X4","X5","X6","X7","X8","X9","X10", "X11","X12","X13","X14","X15","X16", "X17", "X18"), 
               cols =  c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"), 
               ncols = 5,
               pt.size = 1.1,
               sampleids = 5,
               show.sb = FALSE,
               value.scale = "all")

```


```{r, fig.height=10, fig.width=15}

STSC.raw <- GetAssayData(se.pp, assay = "STSC")
mesophyll <- colSums(STSC.raw[c(1, 2, 3, 5, 6,7,8,9,11,12,14),])
vascular <- colSums(STSC.raw[c(4,10,18,15),])
epidermis <- STSC.raw[13,]
guard <- STSC.raw[16,]
hydathode <- STSC.raw[17,]

summed <- t(data.frame(mesophyll,vascular,epidermis,guard,hydathode))


se.pp[['STSCSUM']] <- CreateAssayObject(counts = summed)
DefaultAssay(se.pp) <- "STSCSUM"

plots1 <- c()
for (i in 1:length(section_id)){
  plots1[[i]] =FeatureOverlay(object = SubsetSTData(se.pp, 
                                                   spots = colnames(se.pp)[se.pp$section_id == infoTable$section_id[[i]] & se.pp@assays$STSCSUM@counts["mesophyll",] >= 0.1]),
                             features = "mesophyll",
                             pt.size = 1.1,
                             show.sb = F,
                             cols = c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots2 <- c()
for (i in 1:length(section_id)){
  plots2[[i]] =FeatureOverlay(object = SubsetSTData(se.pp, 
                                                   spots = colnames(se.pp)[se.pp$section_id == infoTable$section_id[[i]] & se.pp@assays$STSCSUM@counts["vascular",] >= 0.1]),
                             features = "vascular",
                             pt.size = 1.1,
                             show.sb = F,
                             cols = c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots3 <- c()
for (i in 1:length(section_id)){
  plots3[[i]] =FeatureOverlay(object = SubsetSTData(se.pp, 
                                                   spots = colnames(se.pp)[se.pp$section_id == infoTable$section_id[[i]] & se.pp@assays$STSCSUM@counts["epidermis",] >= 0.1]),
                             features = "epidermis",
                             pt.size = 1.1,
                             show.sb = F,
                             cols = c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                             dark.theme = F,
                             pt.alpha = 0.8)
}

plots4 <- c()
for (i in 1:length(section_id)){
  plots4[[i]] =FeatureOverlay(object = SubsetSTData(se.pp, 
                                                   spots = colnames(se.pp)[se.pp$section_id == infoTable$section_id[[i]] & se.pp@assays$STSCSUM@counts["guard",] >= 0.1]),
                             features = "guard",
                             pt.size = 1.1,
                             show.sb = F,
                             cols = c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                             dark.theme = F,
                             pt.alpha = 0.8)
}
plots5 <- c()
for (i in 1:length(section_id)){
  plots5[[i]] =FeatureOverlay(object = SubsetSTData(se.pp, 
                                                   spots = colnames(se.pp)[se.pp$section_id == infoTable$section_id[[i]] & se.pp@assays$STSCSUM@counts["hydathode",] >= 0.1]),
                             features = "hydathode",
                             pt.size = 1.1,
                             show.sb = F,
                             cols = c("#fffce8", "#B8B42D", "#B8B42D", "3e363f", "3e363f", "697a21", "697a21"),
                             dark.theme = F,
                             pt.alpha = 0.8)
}

```
##Subset the clusters
```{r, fig.height=6, fig.width=15}

se1 <- SubsetSTData(se.pp, spots = colnames(se.pp)[se.pp$seurat_clusters == "0"])
se2 <- SubsetSTData(se.pp, spots = colnames(se.pp)[se.pp$seurat_clusters == "1"])
se3 <- SubsetSTData(se.pp, spots = colnames(se.pp)[se.pp$seurat_clusters == "2"])
se4 <- SubsetSTData(se.pp, spots = colnames(se.pp)[se.pp$seurat_clusters == "3"])
se5 <- SubsetSTData(se.pp, spots = colnames(se.pp)[se.pp$seurat_clusters == "4"])

cl1Data <- as.data.frame(GetAssayData(se1, assay = "STSCSUM"))
cl2Data <- as.data.frame(GetAssayData(se2, assay = "STSCSUM"))
cl3Data <- as.data.frame(GetAssayData(se3, assay = "STSCSUM"))
cl4Data <- as.data.frame(GetAssayData(se4, assay = "STSCSUM"))
cl5Data <- as.data.frame(GetAssayData(se5, assay = "STSCSUM"))



pheatmap(cl1Data, show_colnames = F, color=colorRampPalette(c("grey", "dark green"))(50))

pheatmap(cl2Data, show_colnames = F, color=colorRampPalette(c("grey", "dark green"))(50))

pheatmap(cl3Data, show_colnames = F, color=colorRampPalette(c("grey", "dark green"))(50))

pheatmap(cl4Data, show_colnames = F, color=colorRampPalette(c("grey", "dark green"))(50))

pheatmap(cl5Data, show_colnames = F, color=colorRampPalette(c("grey", "dark green"))(50))

```


```{r}
writeLines(capture.output(sessionInfo()), "Arabidopsis/10_45_45_all_samples_R/sessionInfo.txt")

```

THE END